### 2019-04-12 Bucketing/sizes.tex
More bucketing stuff

Five definitions appeared only in failing samples, these are named
\texttt{mult2}, \texttt{qexp}, \texttt{op}, \texttt{mul3acc} and \texttt{mul3};
their definitions appear in Figure~\ref{fig:faildefs}. All of these are
functions of natural numbers (\texttt{Nat}). The \texttt{mul3} and
\texttt{mul3acc} functions are rather pathological implementations of
multiplication with an accumulator parameter, involving many (non-tail)
recursive calls which can cause space leaks (or stack overflow, in languages
with eager evaluation). This is likely the reason \textsc{QuickSpec} performs
poorly with these definitions in the input.

The \texttt{op}, \texttt{mult2} and \texttt{qexp} functions are less easily
dismissed since they only recurse in tail position; the latter are in fact
standard definitions of multiplication and exponentiation, respectively. One
potential problem with these definitions, when translated into Haskell for
\textsc{QuickSpec}, is that their accumulating parameter will be lazy by
default, and hence cause space leaks.

@@

  \texttt{
    (define-fun-rec op ((x Nat) (y Nat) (z Nat) (x2 Nat)) Nat
      (match x
        (case Z
          (match z
            (case  Z x2)
            (case (S x3)
              (op Z y x3 (S x2)))))
        (case (S x4)
          (match z
            (case Z
              (op x4 y y x2))
            (case (S c)
              (op x y c (S x2)))))))
  }

@@

>>>>

Five definitions appeared only in failing samples, these are named
\texttt{mul3}, \texttt{mul3acc}, \texttt{mult2}, \texttt{op} and \texttt{qexp};
their definitions appear in Figure~\ref{fig:faildefs}. All of these are
functions of Peano-encoded natural numbers (\texttt{Nat}), and they cause
exploration to time out by generating very large outputs.

\texttt{mul3} and \texttt{mul3acc} are rather pathological implementations of
multiplication with an accumulator parameter, with many (non-tail) recursive
calls. \texttt{mult2} and \texttt{qexp} are standard tail-recursive definitions
of multiplication and exponentiation, respectively. The \texttt{op} function
seems to exist only for testing purposes, appearing in files named
\texttt{weird_nat_op} which assert its commutativity and associativity.

Exploring each of these functions on its own does not require much memory, since
Haskell generates the output lazily. However, comparing such large numbers for
equality takes a lot of CPU time as the \texttt{S} constructors are successively
unwrapped from each side, and this is why the timeout is reached. We confirmed
this hypothesis by exploring with a custom data generator which only generates
the values \texttt{Z}, \texttt{S Z} and \texttt{S (S Z)} (0, 1 and 2); this
caused the exploration to finish quickly. Other interventions, like making the
accumulator arguments strict (to prevent space leaks), did not prevent timeouts.

@@

  \texttt{
    (define-fun-rec op ((x Nat) (y Nat) (z Nat) (x2 Nat)) Nat
      (match x
        (case Z
          (match z
            (case Z      x2)
            (case (S x3) (op Z  y x3 (S x2)))))
        (case (S x4)
          (match z
            (case Z      (op x4 y y  x2))
            (case (S c ) (op x  y c  (S x2)))))))
  }

@@

To assess the impact of these problematic definitions, we removed any samples
containing them and repeated our analysis.
